---
title: "Lab 5"
author: "Muhamad Imran Bin Maszeri"
date: "11/9/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Questions
```{r message=FALSE, warning=FALSE}
library(ggplot2)
library(tidyverse)
library(dplyr)
library(lubridate)
library(readxl)

acc <- read.csv("https://raw.githubusercontent.com/xdaiISU/ds202materials/master/hwlabs/fars2017/accident.csv", stringsAsFactors = FALSE)
person <- read.csv("https://raw.githubusercontent.com/xdaiISU/ds202materials/master/hwlabs/fars2017/person.csv", stringsAsFactors = FALSE)

str(acc)
str(person)
```

1. Are there some days of the week where more accidents happen than the others (see FARS manual, use variable DAY_WEEK)?
```{r message=FALSE, warning=FALSE}
acc$DAY_WEEK <- as.factor(acc$DAY_WEEK)
levels(acc$DAY_WEEK) <- c("Sun", "Mon", "Tue", "Wed", "Thu", "Fri", "Sat")

acc %>%
  group_by(DAY_WEEK) %>%
  summarise(total_accidents=n()) %>%
  arrange(desc(total_accidents))

acc %>%
  ggplot(aes(DAY_WEEK)) + geom_bar() + labs(x='Day', y='Number of Accidents', title='Number of Accidents in Each Day')
```

Answer: More accidents happen in Saturday and Sunday compared to weekdays

2. Create a data frame containing the persons who are fatally hurt in the accidents (look up variable INJ_SEV)
```{r message=FALSE, warning=FALSE}
fatal <- person %>%
  filter(INJ_SEV==4)
str(fatal)

```

3. Create a data frame containing the most dangerous vehicle make in each state. The number of persons fatally hit in the vehicle make is used to assess the (non-)safety of a make. Make sure to handle the missing values appropriately. (look up variable MAKE)
```{r message=FALSE, warning=FALSE}
danger_vehicle <- person %>%
  filter(INJ_SEV==4) %>%
  group_by(STATE, MAKE) %>%
  drop_na() %>%
  summarise(danger=n()) %>%
  group_by(STATE) %>%
  filter(danger==max(danger))

danger_vehicle <- danger_vehicle %>%
  filter(!((STATE==8 & MAKE==37) | (STATE==34 & MAKE==37) | (STATE==50 & MAKE==37)))

danger_vehicle

```

4. Create a map, and label each state with the most dangerous vehicle. Discuss the definition of the most dangerous vehicle, and what you find from the map. (Hint: Read the description for the STATE and COUNTY columns in the FARS manual. The state & county codes are Geographic Locator Codes (GLCs) from the General Services Administration’s (GSA) publication. Use readxl::read_xlsx to read in the GLCs.)
```{r message=FALSE, warning=FALSE}
GLC_USA <- read_xlsx("FRPP_GLC_-_United_StatesSep292020.xlsx")
GLC_USA <- GLC_USA %>%
  mutate(STATE=as.numeric(`State Code`)) %>%
  mutate(COUNTY=as.numeric(`County Code`)) %>%
  mutate(region=tolower(`State Name`)) %>%
  select(!c(`State Code`, `County Code`, `State Name`))

GLC_dat <- data.frame(GLC_USA)

USA_states <- map_data("state")

str(GLC_dat)
str(USA_states)

danger_vehicle <- data.frame(danger_vehicle)

factor(danger_vehicle$MAKE)
danger_vehicle$MAKE <- as.factor(danger_vehicle$MAKE)
levels(danger_vehicle$MAKE) <- c("Ford", "Chevrolet", "Honda", "Toyota")

danger_state <- danger_vehicle %>%
  left_join(GLC_dat, by='STATE', 'COUNTY') %>%
  left_join(USA_states, by='region')

dan_text <- danger_state %>%
  group_by(STATE, MAKE) %>%
  summarise(long=mean(long), lat=mean(lat))

USA_states %>%
  ggplot(aes(x=long, y=lat)) + geom_path(aes(group=group)) + geom_path(aes(group=group)) + geom_text(data=dan_text, aes(x=long, y=lat, label=MAKE), size=2.7) + coord_map() + labs(title='Map of Top Car Maker that Have the Most Accidents')
```

Answer: The most dangerous vehicle in each state is the one with the highest accident in that state. If the accident is similar, the vehicle type is chosen from ascending numericals. Common cars that have accidents in the Midwest is the Chevrolet. The East and West have more Ford cars that are accident.

5. Join the accident and person table (work out which variable(s) to use)
```{r message=FALSE, warning=FALSE}
join_dat <- acc %>%
  inner_join(person, by='ST_CASE')

str(join_dat)

```

6. Tally the number of accidents by day of the week (DAY_WEEK), hour of the day (HOUR) and gender (SEX). Visualize the results.
```{r message=FALSE, warning=FALSE}
join_dat %>%
  group_by(DAY_WEEK) %>%
  tally() %>%
  ggplot(aes(x=DAY_WEEK, y=n)) + geom_bar(stat = 'identity') + labs(x='Day', y='Number of Accidents', title='Number of Accidents in Each Day')

join_dat$SEX <- as.factor(join_dat$SEX)
levels(join_dat$SEX) <- c("Male", "Female", "Not Reported", "Unknown")
join_dat %>%
  filter(!((SEX=="Not Reported") | (SEX=="Unknown"))) %>%
  group_by(SEX) %>%
  tally() %>%
  ggplot(aes(x=SEX, y=n)) + geom_bar(stat = 'identity') + labs(x='Sex', y='Number of Accidents', title='Number of Accidents by Sex')

join_dat %>%
  filter(!((HOUR.x==99) | (HOUR.x==88))) %>%
  group_by(HOUR.x) %>%
  tally() %>%
  ggplot(aes(x=HOUR.x, y=n)) + geom_bar(stat = 'identity') + labs(x='Hour', y='Number of Accidents', title='Number of Accidents in a 24-hour Period')

```

Answer: Male have more accidents in the weekends during the night.

7. Now plot a choropleth map of the number of deaths on a county level.
```{r message=FALSE, warning=FALSE}
county <- map_data("county")

county_sum <- acc %>%
  group_by(COUNTY) %>%
  summarise(death=sum(FATALS))

county_death <- county_sum %>%
  inner_join(GLC_dat, by='COUNTY') %>%
  mutate(subregion=tolower(`County.Name`)) %>%
  select(!County.Name) %>%
  inner_join(county, by='subregion')


str(county_death)

county %>%
  ggplot(aes(x=long, y=lat)) + geom_polygon(aes(group=group)) + geom_polygon(data=county_death, aes(x=long, y=lat, group=group, fill=death)) + coord_map() + labs(title='Map of Concentration of Death in Accidents')
```

Answer: There is a higher rate of accident in the West Coast than any other area.

8. Is summer or winter more dangerous? Does this depend on states? Explore and explain.
```{r message=FALSE, warning=FALSE}
summer <- acc %>%
  filter(MONTH==6 | MONTH==7 | MONTH==8)

summer %>% summarise(sum=n())

summer_state <- summer %>%
  group_by(STATE) %>%
  summarise(total=n()) %>%
  arrange(desc(total)) %>%
  head(5)

summer_state

winter <- acc %>%
  filter(MONTH==1 | MONTH==2 | MONTH==12)

winter %>% summarise(sum=n())

winter_state <- winter %>%
  group_by(STATE) %>%
  summarise(total=n()) %>%
  arrange(desc(total)) %>%
  head(5)

winter_state
```

Answer: Summer is more dangerous with more accidents with 9205 accidents, and Texas and California are both at the top in both seasons, so it does not matter whether any season happen, these 2 states will be on the top accidents.